---
title: "Data Tidying"
author: "AB"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# load packages  
library(dplyr)  
library(tidyr) 
# Produces warning about masking functions with same names in different packages. R assumes most recently added package first. Use 'package_name::function_name(...)' to specify that the other package's function should be used. Add warning = FALSE and message = FALSE to suppress warnings. 
```

# Data Cleaning 

Read in the data.

```{r}
catch <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", 
                  stringsAsFactors = FALSE)  
head(catch)  
```


Keyboard shortcut for pipe operator cmd + shift + m  

```{r}
catch_long <- catch %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  gather(key = "species", value = "catch", -Year, -Region)  
head(catch_long)
```
Select to remove unnecessary columns (marginal sum and notes column). Gather to reformat from wide to long format. Collected species into one species column while keeping Region and Year fields intact. 
Data tidying in progress...  

```{r, eval = F, echo = F}
### test to find the bad data value
test_catch <- as.integer(catch_long$catch)  # try to convert to integer

i <- which(is.na(test_catch)==TRUE)  # find which row(s) have a NA in test_catch  
catch_long[i, ]  # returns row i (i is the index) and all columns
```
Check to find erroneous character value in catch column.   

```{r}
catch_cleaned <- catch_long %>% 
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>%
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)  

head(catch_cleaned)
```

Renamed the catch column to catch_thousands to reflect data units. Corrected character data to integer by replacing the data and converting column to integer.  


# Split-Apply-Combine  

Calculate mean catch by species and year

```{r}
catch_total <- catch_cleaned %>% 
  group_by(species, Year) %>% 
  summarize(catch_mean = mean(catch), 
            n_obs = n())

head(catch_total)  
```

Use filter and arrange



```{r}
catch_chinook <- catch_cleaned %>% 
  #filter(species == "Chinook" | species == "Pink")  # pipe == logical OR
  arrange(-Year) # sort in descending order

head(catch_chinook)
```

What do region codes mean? Join with region table to add columns

# Joins  

Read in Region definition table

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", 
                            method = "libcurl"),
                        stringsAsFactors = FALSE)

head(region_defs)
```

Clean table, removing columns we don't need.

```{r}
region_clean <- region_defs %>% 
  select(code, mgmtArea)

head(region_clean)
```

Now that data are clean, join...

```{r}
catch_joined <- left_join(catch_cleaned, region_clean, 
                          by = c("Region" = "code"))

head(catch_joined)
```

