---
title: "Data Visualization"
author: "AB"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r, echo = F, warning = F, message = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)
```

# Escapement Data

Escapement data are sourced from: 
Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. https://doi.org/10.5063/F1S46Q6M  


```{r, echo = F, results = F}
# echo and results = F makes code run completely silently
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)  

#head(esc)
```

We are interested in annual regional counts of salmon; thus, we need to:  

* 'separate' the date into year - month - day  
* 'group_by' and 'summarize' to calculate the sum over unique region-year-species combinations  
* 'filter' for salmon species  

```{r, echo = F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Chum", "Pink", "Sockeye", "Coho"))
```

```{r, echo = F}
datatable(annual_esc)
```

# Static Plots

```{r, echo = F}
# create theme object to use for plots
my_theme <- theme_light() +
  theme(plot.title = element_text(hjust = 0.5))
```


Escapement for salmon species by region

```{r, echo = FALSE}
# aes call should operate on columns in the data frame
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col() + 
  ggtitle("Salmon Escapement by Region") + 
  ylab("Escapement (number of fish)") +
  my_theme
```

A timeseries of Kodiak escapement by species

```{r, echo = F}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
       mapping = aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() + 
  ylab("Escapement (number of fish)") + 
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
  
```

A timeseries of escapement by species for each regions

```{r, echo = F, fig.height = 10, fig.width= 8}
ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) + 
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free", ncol = 2) +  
  # create plots of individual regions; the tilde is used for linear models, so region is the dependent var in this example 
  # "free_y", "free_x", or "free" allows the axis to scale freely for each subplot
  my_theme

#ggsave("region_escapement.png")  #save the plot as a png in the top-level dir; default saves last plot - save ggplot object to var and add "plot = gg_ob_name" to ggsacve 
```


# Interactive Map

Make a map using leaflet 

* find unique locations with lat/lon values  
* clean to remove observations that are missing a lat or long value

```{r, echo = F}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()  # remove rows with a NA value in any column  
#datatable(locations)  
```

```{r, echo = F}
leaflet(locations) %>% 
  addTiles() %>% # brings in a basemap
  addMarkers(~ Longitude, ~ Latitude, popup = ~ Location)  
```

