---
title: "read_netCDF"
author: "AB"
date: "1/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in a netcdf and make plots

```{r}
library(ncdf4)
library(ggplot2)
library(dplyr)
library(tidyr)
```

# Read in data from a file

Data are from: Fiamma Straneo. 2019. Temperature and salinity profiles adjacent to a tidewater glacier in Sarqardleq Fjord, West Greenland, collected during July 2013. Arctic Data Center. https://doi.org/10.18739/A2B853H78.

```{r}
nc <- nc_open("../data/WG2013CTD.nc")
```

```{r}
print(nc)
```

```{r} 
#list out variable names 
vars <- attributes(nc$var)
vars
```

```{r}
#grab a variable
salinity <- ncvar_get(nc, "sal")
time <- ncvar_get(nc, "time")
depth <- ncvar_get(nc, "z")
```

These ^ data are in a matrix and are not regularly gridded.

# Reformat data

Convert times from matlab time to ISO

```{r}
time <- as.POSIXct((time + 719529)*86400, origin = "1970-01-01", tz = "UTC")
```

Coerce data into a data frame

```{r}
salinity_data <- as.data.frame(salinity)

# Assign column names to depth values
names(salinity_data) <- as.character(depth)


# reshape data and add time

salinity_data_long <- salinity_data %>% 
  mutate(time = time) %>% 
  gather(key = "depth", value = "salinity", -time) %>% 
  mutate(depth = as.numeric(depth))  

```

# Make some plots

```{r}
ggplot(salinity_data_long, aes(x = salinity, y = depth, color = time)) +
    geom_line(size = 0.1) +
    scale_y_reverse()
    #theme_bw() +
    #ylab("Depth (m)") +
    #xlab("") +
    #scale_fill_continuous(low = "gray", high = "red", name = "Salinity (psu)")
```

